Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications

### 2012-07-23 doc/asplos13/relatedwork.tex
added todos to related
another pass on related work. still needs generality claims

\paragraph{Comparable frameworks}
Distributed graph processing frameworks like Pregel \cite{pregel:2010} and
Distributed GraphLab \cite{distgraphlab:vldb12} share similar goals. Pregel
adopts a bulk-synchronous parallel (BSP) execution model, which makes it
inefficient on workloads that could prioritize vertices. GraphLab, on the
other hand, schedules vertex computations individually, allowing
prioritization, which gives faster convergence in a variety of iterative
algorithms.\TODO{GraphLab is not general. Myers will expand on that here. Make
a strong point about generality.} Grappa also supports dynamic parallelism
with asynchronous execution, but parallelism is expressed as tasks or loop
iterations.\TODO{Should RAMcloud go here?}.

\paragraph{Global memory}
Grappa implements a software distributed shared memory DSM system. Many
traditional software DSM systems are page based~\cite{Treadmarks,munin} and
aim to hide the fact that they are built in software from applications by
exploiting the processor's paging mechanisms. Instead Grappa, like other
partitioned global address space (PGAS) models, implements its DSM at the
language, rather than system level. Languages such as Chapel \cite{Chapel},
X10 \cite{X10}, and UPC \cite{UPC} make accesses to shared structures look
like normal memory references. As we describe later, Grappa chooses a middle
ground, where global addresses are explicit in the API and local accesses are
emitted conventionally by the compiler. Similar to DSM systems Grappa provides
a caching and coherence model to ensure high performance and consistent
results. Cache operations are exposed to in the API, however, allowing
software to optimize their usage.

>>>>

\paragraph{Comparable frameworks} Distributed graph processing frameworks like
Pregel \cite{pregel:2010} and Distributed GraphLab \cite{distgraphlab:vldb12}
share similar goals as Grappa. Pregel adopts a bulk-synchronous parallel (BSP)
execution model, which makes it inefficient on workloads that could prioritize
vertices. GraphLab, on the other hand, schedules vertex computations
individually, allowing prioritization, which gives faster convergence in a
variety of iterative algorithms.\TODO{GraphLab is not general. Myers will
expand on that here. Make a strong point about generality.} Grappa also
supports dynamic parallelism with asynchronous execution, but parallelism is
expressed as tasks or loop iterations.\TODO{Should RAMcloud go here?}.

\paragraph{Global memory} Grappa includes a custom implementation of a
software distributed shared memory (DSM) system. Many traditional software DSM
systems are page based~\cite{Treadmarks,munin} and aim to hide the fact that
they are built in software from applications by exploiting the processor's
paging mechanisms, therefore relying heavily on locality. Instead, Grappa,
like other partitioned global address space (PGAS) models, implements its DSM
at the language, rather than system level. Languages such as Chapel
\cite{Chapel}, X10 \cite{X10}, and UPC \cite{UPC} make accesses to shared
structures look like normal memory references. As we describe later, Grappa
chooses a middle ground, where global addresses are explicit in the API and
local accesses are emitted conventionally by the compiler. Similar to DSM
systems, Grappa provides a consistency model that ensure high performance and
with good programmability properties. % Caching is Cache operations are explicit in the  exposed to in the
% API, however, allowing software to optimize their usage.

We decided to build a custom distributed shared memory implementation, as
opposed to using an existing implementation, for two key reasons: (1) in order
to support the large number of tasks required for latency tolerance, the
implementation of the global memory system is tightly coupled to the task
scheduler; and (2) since commodity networks only do efficient RDMA or ordinary
transmission of small messages, the implementation of the global memory system
must make use of the network message aggregator.

%
